The four postulates of quantum mechanics are three

### 2020-11-20 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
Changes with Lorenzo

%ATTENZIONE: PER COMPILARE BISOGNA FARE A MANO
%dvipdf tensor.dvi ;mv tensor.pdf temp.pdf 
%Dovremmo forse aggiungere due parole sulle C* algebras: non sono stateintrodotte proprio per evitare il tensor product in infinitedimensional systems?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The four postulates of quantum mechanics are three
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[aps,prl,amsmath,amssymb,twocolumn,nofootinbib]{revtex4}
%\documentclass[twocolumn,aps,showpacs,prl,groupedaddress]{revtex4}
%\usepackage{amssymb,stackengine}
\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epsfig,amssymb,amsmath,amsthm}
%\usepackage[active]{srcltx}
%\usepackage[hypertex,linkcolor=red]{hyperref}

@@

\def\>{\rangle}
\def\<{\langle}
\def\ca{_{\cal A}}
\def\cb{_{\cal B}}
\def\cc{_{\cal C}}
%\def\comment#1{}
\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
\def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
\def\labell#1{\label{#1}}
%\def\labell#1{\label{#1}{\mbox{{\tiny #1}}}}
%\def\section#1{{\par\em #1:--- }}
\def\togli#1{}
\def\sh{\mbox{sh}}
\def\iden{\openone}
\begin{document}

@@

The tensor product postulate does not appear in all axiomatizations of
quantum mechanics: it has even been called ``postulate 0'' in some
literature \cite{zurek}. A widespread belief is that it is a direct
consequence of the superposition principle, and it is hence not a
necessary axiom. {\em This belief is mistaken}: the superposition
principle is encoded into the quantum axioms by requiring that the
state space is a {\em linear} vector space. This is, by itself,
insufficient to single out the tensor product, as other linear
products of linear spaces exist, such as the direct product, the exterior/wedge roduct or the direct sum of vector spaces, which is used
	in classical mechanics to combine state spaces of linear systems. These are all maps from linear spaces to linear spaces but they differ in how the linearity of one is mapped to the linearity of the other.\footnote{For example, in the tensor product $a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct product $a \times (b+c) = a \times b + 0 \times c$ where $0$ is the zero vector. Also, in the tensor product $r * (a \otimes b) = (r*a) \otimes b = a \otimes (r*b)$ while in the direct product $r * (a \times b) = r*a \times r*b$.}

	This belief may have arisen from the seminal book of Dirac
	\cite{diracbook}, who introduces tensor products (Chap.~20) by
	appealing to linearity. However, he adds the seemingly innocuous
	request that the product among spaces be distributive (rather,
	bilinear), which is equivalent to postulating tensor products (or
	linear functions of them). This is not an innocuous request. For
	example it does not hold where the composite vector space of two
	linear spaces is described by the direct product, e.g.~in classical
	mechanics, for two strings of a guitar: it is not distributive.
	[General classical systems, not only linear ones, are also composed
	through the direct product.] Of course, Dirac is not constructing an
	axiomatic formulation, so his `sleight of hand' can be forgiven. In
	contrast, von Neumann (\cite{vonneumannbook} Chap.~VI.2, also
	\cite{jauch}) introduces tensor products by noticing that this is a
	natural choice in the position representation of wave mechanics (where
	they were introduced in \cite{weyl,epr}), and then {\em explicitly
		postulates} them in general: ``This rule of transformation is
	correct in any case for the coordinate and momentum operators [...]
	and it conforms with the [observable axiom and its linearity
	principles], we therefore postulate them generally.''
	\cite{vonneumannbook}.  More mathematical or conceptually-oriented
	modern formulations (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang})
	introduce this postulate explicitly.  An interesting alternative is
	provided in \cite{ballentinebook,ballentinepaper}: after introducing
	tensor products, Ballentine verifies a posteriori that they give the
	correct laws of composition of probabilities. Similarly, Peres uses
	relativistic locality \cite{peres}. While these procedures seemingly
	bypass the need to postulate the tensor product, they do not guarantee
	that this is the {\em only} possible way of introducing composite
	systems in quantum mechanics. In the framework of quantum logic,
	tensor products arise from some additional conditions \cite{matolcsi}
	which (in contrast to what is done here) are not connected to the
	other postulates. Similar approaches appear in \cite{marmo,aerts}
	where tensor products were obtained by specifying additional physical
	or mathematical requirements. In quantum field theory one tends to
	avoid problems connected with tensor products of infinite dimensional
	spaces by focusing on algebraic commutation structures,
	e.g.~\cite{giddins,roos}.  In particular, the recent MIP*=RE result
	\cite{mipre} implies that, in infinite dimensions, the tensor product
	is strictly less computationally powerful than the commutation
	structures, emphasizing the difference among these two structures, at
	least for the infinite-dimensional case. We will consider the
	non-relativistic setting here.

	In this paper we derive the tensor product postulate (which, hence,
	loses its status of postulate) from two other postulates of quantum
	mechanics: the state postulate and the measurement postulate. We start
	from the natural definition of a composite system as the set of two
	(or more) quantum systems. The composite system is therefore
	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else},
	it's state being fully described by all observables and only the observables
	that fully describe its constituents. The first key insight is that first two postulates already assume that, at least in some cases, the probability distribution of the
	measurement outcomes of one system is independent of the other systems' (statistical independence). While we will see specifically how the Born implies this, it should be clear conceptually that we would not be able to talk about a system in the first place if we would not be able to characterize it independently. The second key insight is that, since all measurement
	outcomes and all states of the subsystems plus the composite live in the same probability space, we will automatically have a map $M$ that for a pair of subsystems' states gives us the composite state that corresponds to the statistically independent case. These insights are enough to characterize mathematically the state space of the composite: the linearity given by the Hilbert space, together with the fact that the composite system is fully described by the observables of $A$ and $B$, allows to extend the construction to the composite states that do not satisfy statistical independence (the entangled states). So the work consists of two interrelated efforts: a physical argument that start from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.

	We will mainly focus on
	kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no superselection rules or
	other restrictions to the state space are present: it is possible to
	prepare each subsystem of a composite system in a state that is
	independent of the other systems (preparation independence).  This is
	the only case in which the tensor product can be properly employed
	\cite{susskind,zanardi,zanardilloyd}: the Hilbert space of composite
	systems that have restrictions is {\em not} the tensor product of the
	component spaces, but a subspace of it (e.g.~the anti-symmetric
	subspace for fermions). Typically this is ignored in the literature,
	since the tensor product formalism is very convenient and is often
	used also in these cases, but it is something we need to bring out.

	While conceptually the derivation is straightforward, it requires the
	use of projective spaces, which may be unfamiliar to the reader. This is
	necessary because events in probability space do not correspond to vectors in the Hilbert space, but to subspaces. If $\psi$ is a vector, the state corresponds to the ray $\pj{\psi}$, which is the subspace spanned by $\psi$. Note that ``ray'' can be misleading since $\pj{\psi}$ is really a complex { \em plane }. In the same way, the event $X = x_0$ corresponds to the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$. The map $M$ derived from the probability space is a map between states as events in probability space, and therefore will be a map from rays to rays, which is a map between the projective spaces. The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.

	This is done by showing that preparation independence and
	statistical independence imply three conditions on the map $m$:
	(H1)~totality: the map is defined on all states of the subsystems;
	(H2)~bilinearity: the map is bilinear thanks to the fundamental
	theorem of projective geometry; (H3)~span surjectivity: the span of
	the image of map coincides with the full composite Hilbert space.  We
	then prove that, if these three conditions H1, H2 and H3 hold, then
	the map $m$ is the tensor product, namely the Hilbert space of the
	composite system is a tensor product of the components: the tensor
	product ``postulate'', which hence loses its status of a postulate. An
	overview of all these logical implications is given in
	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
	argument. Refer to the Supplementary Information for a more rigorous
	proof of the \emph{same} argument.

@@

	We start with the axiomatization of quantum mechanics based on the following
	postulates (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state of a
	system is described by a ray $\pj{\psi}$ corresponding to a set of
	non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
	system's observable properties are described by self-adjoint operators
	acting on that space; (b)~The probability that a measurement of a
	property $X$, described by the operator with spectral decomposition
	$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
	$x$ given that the system is in state $\pj{\psi}$ is
	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2$ (Born rule). (c)~The state
	space of a composite system is given by the tensor product of the
	spaces of the component systems; (d)~The time evolution of an isolated
	system is described by a unitary operator acting on a vector
	representing the system state, $|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or,
	equivalently, by the Schr\"odinger equation. The rest of quantum
	theory can be derived from these axioms. While some axiomatizations
	introduce further postulates, we will be using only (a) and (b) to
	derive (c), so the above are sufficient for our aims.

	\togli{This axiomatization implicitly contains a definition of
		``quantum system'' which is crucial for what follows, so we need to
		clarify the assumptions that it contains. We will use the following
		definition for a quantum
		system\togli{$\stackon[1pt]={\mbox{\tiny
					def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree
			of freedom with $d$ (possibly discrete, or continuous, infinite)
			mutually exclusive (commuting) values for each of its properties.
			Its mathematical description is through a Hilbert space of
			dimension $d$ which contains all the states that describe the
			values of its possible properties. In accordance with the
			postulate (a), these values correspond to a basis of the space,
			given by the eigenvectors of the observable corresponding to that
			property''}. \commentg{We may have to revise to be more clear.
			Where is this used?} } As discussed, we limit ourselves to
	kinemati\-cal\-ly-inde\-pen\-dent systems, where all state vectors
	$|\psi\>$ in the system's Hilbert space $\cal H$ describe a valid
	state, {\em unconditioned on anything else}. In particular, we note
	that restrictions due to superselection rules arise either from
	practical (not fundamental) limitations on the actions of the
	experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
	ill-defined quantum systems. For example, in situations where
	indistinguishability plays a role (e.g.~in QFT), one cannot consider
	an electron as a quantum system: in this case the quantum system is
	the field. The electron is an excitation of the field, {\em not} a
	(sub)system. We call this condition ``preparation independence''.  [We
	emphasize that the kinematic independence is inequivalent to dynamical
	independence (or isolation).  Indeed if two systems interact, their
	interaction may lead to dynamical restrictions in the state spaces.
	Here we will not consider dynamical evolution, which is contained in
	postulate (d).]

	The definition of a composite system as containing {\em only} the
	collection of the subsystems means that any preparation of both
	subsystems independently must correspond to the preparation of the
	composite system. Since states are defined by postulate (a) as rays in
	the respective Hilbert spaces, there must exist a map $M : \pj{\cal A}
	\times \pj{\cal B } \to \pj{\cal C}$ that takes a pair of states for
	the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$ represent the
	projective spaces and the Cartesian product is the set of all possible
	pairs) and returns a state in the projective space $\pj{\cal C}$ for
	the composite. To visualize the geometrical meaning of $M$ directly within the Hilbert spaces, given a ray (a complex plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex plane) in $\cal C$.  Our final goal will be to find a map
	$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
	Hilbert spaces $\cal A$, $\cal B$ and $\cal C$ consistently with $M$. Namely,
	$\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign indicates the
	elements in the projective space. Again geometrically, $m$ takes a vector in each of $\cal A$ and $\cal B$, and returns a vector in $\cal C$ and we want this to be consistent with $M$ such that vectors picked from the same rays will return vectors in the same ray. We will prove that the map $m$ is
	the tensor product. We focus on pure states here: the argument can be
	extended to mixed states using standard tools \cite{nielsenchuang}.

	The map $M$ must be injective: as said above, different states of the
	subsystems must correspond, by definition of composite system, to
	different states of the composite. Moreover, preparation independence
	implies that $M$, and hence $m$, must be total maps: each subsystem of
	the composite system can be independently prepared and gives rise to a
	state of the composite (condition H1).  H1 is not sufficient
	to identify the tensor product: by itself it does not even guarantee
	that the map $m$ is linear.

	Postulate (b) contains the connection between quantum mechanics
	and probability theory. It must then implicitly contain the
	axiomatization of probability, e.g.~see
	\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
	probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
	the joint probability of events $a$ and $b$ given $z$ is
	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. This essentially the definition of conditional probability: $p(b|z\wedge a) = \frac{p(a\wedge b|z)}{p(a|z)}$. Two events $a$ and $b$ are defined to be 
	independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$. However, it is more instructive to define independence with $p(b|z\wedge a)=p(b|z)$. This makes it more evident that the probability of $b$ does not depend on conditioning on $a$. If we imagine $a$ and $b$ to be outcomes for particular measurements, we are saying that the probability of observing $b$ is not a function of the value measured by $a$. We can recover the first definition from the second. In fact, $p(b|z\wedge a)=\frac{p(a\wedge b|z)}{p(a|z)} =p(b|z)$ which leads to $p(a\wedge b|z)=p(a|z)\:p(b|z)$.

	We note that the Born
	probability formula in postulate (b) is of the form $p(a|\psi_{\cal A}$ and contains only quantities of the
	system (the pure state $\psi_{\cal A}$ and the observable's eigenstates $a$). Now, suppose that, unknown to us, an additional second system was prepared in a pure state $\psi_{\cal B}$ and consider $p(a|\psi_{\cal A}\wedge \psi_{\cal B})$. In what cases are we allowed to use the Born rule? That is, in what cases can we say $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = |\<a|\psi_{\cal A}\>|^2$? Only in the cases where $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = p(a|\psi_{\cal A}$, where the systems  are statistically independent. In this sense, postulate (b) already implies  ``statistical independence'' between
	different systems if they are prepared independently. Namely, the
	outcome probability of one is independent of any properties of any
	other in the absence of any dynamical coupling and supposing
	independent preparation. [Of course, the dynamics may couple the state
	of the system to other systems so that the outcomes may depend on what
	happens to other systems.  Analogously, one may prepare the system in
	a way that depends on other systems. But we consider independent
	preparations and do not consider dynamics, as it is sufficient for our
	aims.] We can then formalize ``statistical independence'' of
	independently prepared systems by saying that their Born rule
	satisfies:
	\begin{eqnarray}
	p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
		B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
		A})\:p(b|\psi_{\cal B})
	\labell{bindip}\;,
	\end{eqnarray}
	where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
	preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
	$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
	B$ respectively. The first equality in equation \eqref{bindip} embodies the
	definition of statistical independence, the second follows from the
	form of the Born rule for independently prepared systems (the
	probability depends only on the state and observables of each system
	on its own).

>>>>

%ATTENZIONE: PER COMPILARE BISOGNA FARE A MANO
%dvipdf tensor.dvi ;mv tensor.pdf temp.pdf 
%Dovremmo forse aggiungere due parole sulle C* algebras: non sono stateintrodotte proprio per evitare il tensor product in infinitedimensional systems?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The four postulates of quantum mechanics are three
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[aps,prl,amsmath,amssymb,twocolumn]{revtex4}
%\documentclass[twocolumn,aps,showpacs,prl,groupedaddress]{revtex4}
%\usepackage{amssymb,stackengine}
\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epsfig,amssymb,amsmath,amsthm}
%\usepackage[active]{srcltx}
%\usepackage[hypertex,linkcolor=red]{hyperref}

@@

\def\>{\rangle}
\def\<{\langle}
\def\ca{_{\cal A}}
\def\cb{_{\cal B}}
\def\cc{_{\cal C}}
\def\comment#1{}
%\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
\def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
\def\labell#1{\label{#1}}
%\def\labell#1{\label{#1}{\mbox{{\tiny #1}}}}
%\def\section#1{{\par\em #1:--- }}
\def\togli#1{}
\def\sh{\mbox{sh}}
\def\iden{\openone}
\begin{document}

@@

	In this paper we derive the tensor product postulate (which, hence,
	loses its status of postulate) from two other postulates of quantum
	mechanics: the state postulate and the measurement postulate.
	The tensor product postulate does not appear in all axiomatizations of
	quantum mechanics: it has even been called ``postulate 0'' in some
	literature \cite{zurek}. A widespread belief is that it is a direct
	consequence of the superposition principle, and it is hence not a
	necessary axiom. {\em This belief is mistaken}: the superposition
	principle is encoded into the quantum axioms by requiring that the
	state space is a {\em linear} vector space. This is, by itself,
	insufficient to single out the tensor product, as other linear
	products of linear spaces exist, such as the direct product, the exterior/wedge product or the direct sum of vector spaces, which is used
	in classical mechanics to combine state spaces of linear systems. These are all maps from linear spaces to linear spaces but they differ in how the linearity of one is mapped to the linearity of the other.\footnote{For example, in the tensor product $a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct product $a \times (b+c) = a \times b + 0 \times c$ where $0$ is the zero vector. Also, in the tensor product $r  (a \otimes b) = (r a) \otimes b = a \otimes (r b)$ while in the direct product $r  (a \times b) = r a \times r b$, where $r$ is a scalar.}
	This belief may have arisen from the seminal book of Dirac
	\cite{diracbook}, who introduces tensor products (Chap.~20) by
	appealing to linearity. However, he adds the seemingly innocuous
	request that the product among spaces be distributive (rather,
	bilinear), which is equivalent to postulating tensor products (or
	linear functions of them). It is not an innocuous request. For
	example it does not hold where the composite vector space of two
	linear spaces is described by the direct product, e.g.~in classical
	mechanics, for two strings of a guitar: it is not distributive.
	[General classical systems, not only linear ones, are also composed
	through the direct product.] Of course, Dirac is not constructing an
	axiomatic formulation, so his `sleight of hand' can be forgiven. In
	contrast, von Neumann (\cite{vonneumannbook} Chap.~VI.2, also
	\cite{jauch}) introduces tensor products by noticing that this is a
	natural choice in the position representation of wave mechanics (where
	they were introduced in \cite{weyl,epr}), and then {\em explicitly
		postulates} them in general: ``This rule of transformation is
	correct in any case for the coordinate and momentum operators [...]
	and it conforms with the [observable axiom and its linearity
	principles], we therefore postulate them generally.''
	\cite{vonneumannbook}.  More mathematical or conceptually-oriented
	modern formulations (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang})
	introduce this postulate explicitly.  An interesting alternative is
	provided in \cite{ballentinebook,ballentinepaper}: after introducing
	tensor products, Ballentine verifies a posteriori that they give the
	correct laws of composition of probabilities. Similarly, Peres uses
	relativistic locality \cite{peres}. While these procedures seemingly
	bypass the need to postulate the tensor product, they do not guarantee
	that this is the {\em only} possible way of introducing composite
	systems in quantum mechanics. In the framework of quantum logic,
	tensor products arise from some additional conditions \cite{matolcsi}
	which (in contrast to what is done here) are not connected to the
	other postulates. Similar approaches appear in \cite{marmo,aerts}
	where tensor products were obtained by specifying additional physical
	or mathematical requirements. \comment{ {\em Il resto della frase da spostare in supp material} In quantum field theory one tends to
		avoid problems connected with tensor products of infinite dimensional
		spaces by focusing on algebraic commutation structures,
		e.g.~\cite{giddins,roos}.  In particular, the recent MIP*=RE result
		\cite{mipre} implies that, in infinite dimensions, the tensor product
		is strictly less computationally powerful than the commutation
		structures, emphasizing the difference among these two structures, at
		least for the infinite-dimensional case. We will consider the
		non-relativistic setting here.}

	We start
	from the natural definition of a composite system as the set of two
	(or more) quantum systems. The composite system is therefore
	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else}\comment{da mettere altrove:, namely, 
		its state is fully described by the probability amplitudes of the system's observables and nothing else}. The first key insight is that the first two postulates of quantum theory (introduced below) already assume that, at least in some cases, the probability distribution of the
	measurement outcomes of one system is independent of the other systems' (statistical independence). We cannot even talk about a system in the first place if we cannot characterize it independently. The second key insight is that, since all measurement
	outcomes and all states of the subsystems plus the composite live in the same probability space, we will automatically have a map $M$ that for a pair of subsystems' states gives us the composite state that corresponds to the statistically independent case. These insights are enough to characterize mathematically the state space of the composite: the linearity given by the Hilbert space, together with the fact that the composite system is fully described by the observables of $A$ and $B$, allows to extend the construction to the composite states that do not satisfy statistical independence (the entangled states). So the work consists of two interrelated efforts: a physical argument that starts from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.

	\comment{mettere nel supplementary:	We will mainly focus on
		kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no superselection rules or
		other restrictions to the state space are present: it is possible to
		prepare each subsystem of a composite system in a state that is
		independent of the other systems (preparation independence).  This is
		the only case in which the tensor product can be properly employed: the Hilbert space of composite
		systems that have restrictions is {\em not} the tensor product of the
		component spaces, but a subspace of it (e.g.~the anti-symmetric
		subspace for fermions). Typically this is ignored in the literature,
		since the tensor product formalism is very convenient and is often
		used also in these cases, and superselection rules are typically avoidable \cite{susskind,zanardi,zanardilloyd}. }

	TODO: pensarci con calma: riscrivere il paragrafo seguente, magari partendo da un esempio e introducendo il concetto seguente:
	Frase in cui diciamo che abbiamo bisogno di projective spaces perche' i vettori $|\psi\>$ non rappresentano proprieta' (direbbe Lorenzo) o events (o proposizioni) in probability space (direbbe Gabriele) ma sono i sottospazi che lo fanno.

	Since quantum states are rays in Hilbert space, we need projective spaces
	Events in probability space do not correspond to vectors in the Hilbert space, but to subspaces, which means we will need elements of projective geometry. If $\psi$ is a vector, the state corresponds to the ray $\pj{\psi}$, which is the subspace spanned by $\psi$. Note that ``ray'' can be misleading since $\pj{\psi}$ is really a complex { \em plane }. In the same way, the event $X = x_0$ corresponds to the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$. The map $M$ derived from the probability space is a map between states as events in probability space, and therefore will be a map from rays to rays, which is a map between the projective spaces. The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.

	This is done by showing that preparation independence and
	statistical independence imply three conditions on the map $m$:
	(H1)~totality: the map is defined on all states of the subsystems;
	(H2)~bilinearity: the map is bilinear thanks to the fundamental
	theorem of projective geometry; (H3)~span surjectivity: the span of
	the image of map coincides with the full composite Hilbert space.  We
	then prove that, if these three conditions H1, H2 and H3 hold, then
	the map $m$ is the tensor product, namely the Hilbert space of the
	composite system is a tensor product of the components: the tensor
	product ``postulate'', which hence loses its status of a postulate. An
	overview of all these logical implications is given in
	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
	argument. The supplementary material contains the mathematical details.

@@

	We start with the axiomatization of quantum mechanics based on the following
	postulates (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state of a
	system is described by a ray $\pj{\psi}$ corresponding to a set of
	non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
	system's observable properties are described by self-adjoint operators
	acting on that space; (b)~The probability that a measurement of a
	property $X$, described by the operator with spectral decomposition \comment{Normalizzare qui e in altri posti!}
	$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
	$x$ given that the system is in state $\pj{\psi}$ is
	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2$ (Born rule). (c)~The state
	space of a composite system is given by the tensor product of the
	spaces of the component systems; (d)~The time evolution of an isolated
	system is described by a unitary operator acting on a vector
	representing the system state, $|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or,
	equivalently, by the Schr\"odinger equation. The rest of quantum
	theory can be derived from these axioms. While some axiomatizations
	introduce further postulates, we will be using only (a) and (b) to
	derive (c), so the above are sufficient for our aims.

	\togli{This axiomatization implicitly contains a definition of
		``quantum system'' which is crucial for what follows, so we need to
		clarify the assumptions that it contains. We will use the following
		definition for a quantum
		system\togli{$\stackon[1pt]={\mbox{\tiny
					def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree
			of freedom with $d$ (possibly discrete, or continuous, infinite)
			mutually exclusive (commuting) values for each of its properties.
			Its mathematical description is through a Hilbert space of
			dimension $d$ which contains all the states that describe the
			values of its possible properties. In accordance with the
			postulate (a), these values correspond to a basis of the space,
			given by the eigenvectors of the observable corresponding to that
			property''}. \commentg{We may have to revise to be more clear.
			Where is this used?} } As discussed, we limit ourselves to
	kinemati\-cal\-ly-inde\-pen\-dent systems, where all state vectors
	$|\psi\>$ in the system's Hilbert space $\cal H$ describe a valid
	state, {\em unconditioned on anything else}. In particular, we note
	that restrictions due to superselection rules arise either from
	practical (not fundamental) limitations on the actions of the
	experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
	ill-defined quantum systems. \footnote{For example, in situations where
		indistinguishability plays a role (e.g.~in QFT), one cannot consider
		an electron as a quantum system: in this case the quantum system is
		the field. The electron is an excitation of the field, {\em not} a
		(sub)system.} We call this condition ``preparation independence''.  [We
	emphasize that the kinematic independence is inequivalent to dynamical
	independence (or isolation).  Indeed if two systems interact, their
	interaction may lead to dynamical restrictions in the state spaces.
	Here we will not consider dynamical evolution, which is contained in
	postulate (d).]

	The definition of a composite system as containing {\em only} the
	collection of the subsystems means that any preparation of both
	subsystems independently must correspond to the preparation of the
	composite system. Since states are defined by postulate (a) as rays in
	the respective Hilbert spaces, there must exist a map $M : \pj{\cal A}
	\times \pj{\cal B } \to \pj{\cal C}$ that takes a pair of states for
	the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$ represent the
	projective spaces and the Cartesian product is the set of all possible
	pairs) and returns a state in the projective space $\pj{\cal C}$ for
	the composite. To visualize the geometrical meaning of $M$ directly within the Hilbert spaces, given a ray (a complex plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex plane) in $\cal C$.  Our final goal will be to find a map
	$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
	Hilbert spaces $\cal A$, $\cal B$ and $\cal C$ consistently with $M$. Namely,
	$\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign indicates the
	elements in the projective space. Again geometrically, $m$ takes a vector in each of $\cal A$ and $\cal B$, and returns a vector in $\cal C$ and we want this to be consistent with $M$ such that vectors picked from the same rays will return vectors in the same ray. We will prove that the map $m$ is
	the tensor product. We focus on pure states here: the argument can be
	extended to mixed states using standard tools \cite{ballentine}.

	The map $M$ must be injective: as said above, different states of the
	subsystems must correspond, by definition of composite system, to
	different states of the composite. Moreover, preparation independence
	implies that $M$, and hence $m$, must be total maps (condition H1): each subsystem of
	the composite system can be independently prepared and gives rise to a
	state of the composite.  H1 is not sufficient
	to identify the tensor product: by itself it does not even guarantee
	that the map $m$ is linear.

	Postulate (b) contains the connection between quantum mechanics
	and probability theory. It must then implicitly contain the
	axiomatization of probability, e.g.~see
	\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
	probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
	the joint probability of events $a$ and $b$ given $z$ is
	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Two events $a$ and $b$ are defined to be 
	independent given $z$ if $p(a\wedge b|z)=p(a|z)\:p(b|z)$, which is true if and only if $p(b|z\wedge a)=p(b|z)$. If $a$ and $b$ are measurement outcomes, the probability of $b$ is not a function of $a$.

	Riscrivere questo: scritto cosi', sembra che la regola di Born non si possa applicare sempre....
	We note that the Born
	probability formula in postulate (b) is of the form $p(a|\psi_{\cal A}$ and contains only quantities of the
	system (the pure state $\psi_{\cal A}$ and the observable's eigenstates $a$). Now, suppose that, unknown to us, an additional second system was prepared in a pure state $\psi_{\cal B}$ and consider $p(a|\psi_{\cal A}\wedge \psi_{\cal B})$. In what cases are we allowed to use the Born rule? That is, in what cases can we say $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = |\<a|\psi_{\cal A}\>|^2$? Only in the cases where $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = p(a|\psi_{\cal A}$, where the systems  are statistically independent. In this sense, postulate (b) already implies  ``statistical independence'' between
	different systems if they are prepared independently. Namely, the
	outcome probability of one is independent of any properties of any
	other in the absence of any dynamical coupling and supposing
	independent preparation. [Of course, the dynamics may couple the state
	of the system to other systems so that the outcomes may depend on what
	happens to other systems.  Analogously, one may prepare the system in
	a way that depends on other systems. But we consider independent
	preparations and do not consider dynamics, as it is sufficient for our
	aims.] We can then formalize ``statistical independence'' of
	independently prepared systems by saying that their Born rule
	satisfies:
	\begin{eqnarray}
	p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
		B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
		A})\:p(b|\psi_{\cal B})
	\labell{bindip}\;,
	\end{eqnarray}
	where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
	preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
	$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
	B$ respectively. The first equality in equation \eqref{bindip} embodies the
	definition of statistical independence, the second follows from the
	form of the Born rule for independently prepared systems (the
	probability depends only on the state and observables of each system
	on its own).
